---
title: "Day 1 Examples"
author: "Nicole LaRoche"
date: "11/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Header 1

## Header 2
 
Some **bold**

Include a [link](https://nceas.ucsb.edu)

 * apple
 * banana
 * orange

*italic*

1. this is number 1
    * This is an indent
2. this is number 2 

Here is some R code:

```{r}
x <- 3*4
x
```


This should be inline `echo=FALSE` in the middle of the sentence

Another example of the same thing. This will print the value of x=`r round(x)`

### LaTeX example

$\sum{x^2}$

### Now we are hiding the results of the r code and only giving a result

```{r, results='hide', echo = FALSE}

y=10
z<-x*y
```

The result is `r z`. 

```{r}
weight_kg <- c(55, 25, 12)
mean(weight_kg)
```

# read this data into R environment

Use `read.csv` to read the data

```{r}
bg_chem <- read.csv("BGchem2008data.csv", stringsAsFactors = FALSE)
class(bg_chem)
head(bg_chem)
mean_depth <- mean(bg_chem$CTD_Depth)
```

```{r}
plot(bg_chem$CTD_Depth, bg_chem$CTD_Temperature)
```


