---
title: "Data_Tidy"
author: "Nicole LaRoche"
date: "11/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning =F, message = F}
library(dplyr)
library(tidyr)
```
# Outline

## Read and Clean Data
```{r}
catch_original <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1",
                           stringsAsFactors = FALSE)
head(catch_original)
```

We want to drop the `all` and `notes` columns
```{r}
catch_data <- catch_original %>% 
  #select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)
  select(-All, -notesRegCode)
```

```{r}
summary(catch_data)
```

#### Fix Chinook column
Chinook has a problem. It is listed as a character. 
use `mutate` to fix
```{r}
catch_clean <- catch_data %>% 
  mutate(Chinook = ifelse(Chinook == "I", 1, Chinook)) %>% 
  mutate(Chinook = as.numeric(Chinook))
```

data exploration - find the row that got turned into NA in `Chinook`
```{r}
i<-which(is.na(catch_clean$Chinook))
i
```

```{r}
catch_original[i, ]
```

## Reshape Data

Using `pivot_longer` to condense the species into one column.
```{r}
catch_long<- catch_clean %>% 
  pivot_longer(cols = -c(Region, Year), names_to = "Species", values_to = "Count")
head(catch_long)
```


Using `pivot_wider` to make the species separate again
```{r}
catch_wide<- catch_long %>% 
  pivot_wider(names_from = Year, values_from = Count)
head(catch_wide)
```

`rename` count column to `catch_thousands`

```{r}
catch_long<- catch_long %>% 
  rename(catch_thousands = Count)
```

```{r}
catch_long<- catch_long %>% 
  mutate(catch = catch_thousands*1000) %>% 
  select(-catch_thousands)
head(catch_long)
```

## Summarize data

```{r}
mean_region <- catch_long %>% 
  group_by(Region, Species) %>% 
  summarise(catch_mean= mean(catch), num_obs = n())

mean_region
```

Now calculate the most commonly caught species
First, calculate the total catch by species
```{r}
max_species <- catch_long %>% 
  group_by(Species) %>% 
  summarise(total = mean(catch)*n()) %>% 
  arrange(desc(total))
max_species
```

filter - this isn't done. Come back to it.
```{r}
species_pre_1900 <- catch_long %>% 
  filter(Year<1900) %>% 
  group_by(Species) %>% 
  summarise(total = mean(catch)*n()) %>% 
  arrange(desc(total))
species_pre_1900
```



## Join to region table


Ading second dataset
```{r}
region_defs <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1",
                        stringsAsFactors = FALSE) %>% 
  select(code, mgmtArea)
head(region_defs)
```

Left join - making the combined file with all values from the catch data. if there is something in the regions data, that doesn't have a catch zone associated then it will not appear in this list.
```{r}
catch_joined <- left_join(catch_long, region_defs, by = c("Region"="code"))
head(catch_joined)
```

# Separate and Unite

```{r}
site_codes <- data.frame(site = c("HAW-100", "HAW-101", "OAH-102", "OAH-103", "MAI-100"),
                         stringsAsFactors = F)
```

separate island and site number using `separate`
```{r}
site_codes_split<-site_codes %>% 
  separate(site, into = c("island", "site_num"), sep = "-", remove = F)

site_codes_split
```

now undo what we just did
```{r}
site_codes_combine <- site_codes_split %>% 
  unite(site_code, island, site_num, sep = "_")

site_codes_combine
```

